\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-01-26 draft.tex
extensions section removed, fn on asymmetry added

I will begin by describing the model in detail. Section~\ref{sec:main} then presents the main results. In Section~ \ref{sec:example}, an extended example demonstrates these results as well as the impact of changing the institutional environment. Section~\ref{sec:dis} explores the connection between the model under consideration here and the `Protection for Sale' theoretical framework and includes in-depth discussions of the roles of uncertainty and the separation-of-powers structure. Several extensions of the model are explored in Section~\ref{sec:ext} and Section~\ref{sec:concl} concludes.

@@

I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small. This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries. I discuss an extension to the case in which both legislatures vote to break the agreement at the same time in Section~\ref{sec:asym}.

@@

\subsection{The Trade Agreement}
\label{sec:ta}
The executives choose trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ to maximize Equation \ref{eq:jv}. I call the maximized joint value $MV(\bta)$ and write the division of surplus according to the Nash bargaining solution, which, due to symmetry, can be simplified from its general multiplicative form as
\[
  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
\]
\[
  V_E^*(\bta) - t = W_E^*(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
\]
where expectation operators have been dropped for convenience and $t$ is an intergovernmental transfer.\footnote{While direct monetary transfers have to date rarely been used in practice, it seems appropriate to interpret linked concessions on non-trade issues as indirect transfers (Klimenko, Ramey and Watson (2008); Maggi and Staiger (2011)). Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in the environment of Section~\ref{sec:asym}.} The solution to this system of equations determines the tariffs $\bta$ that will be set in the trade agreement.

@@

The problem then becomes to maximize the following modified version of Equation \ref{eq:jv}:
\begin{equation}
    \left\{ o \cdot B(\bta) + o^* \cdot B^*(\bta) \right\} \bm{W_E}(\btn) + \left\{ 1- o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \bm{W_E}(\bta)  
  \label{eq:jv2}
\end{equation}
where I have written the sum of the home and foreign executives' utilities as $\bm{W_E}(\cdot)$. Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement. The case in which both legislatures can potentially break the agreement is treated in Section~\ref{sec:two}. One can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.

@@

\section{Extensions}
\label{sec:ext}
\begin{comment}
This analysis has been undertaken with the assumption that the import-competing industry is distributed evenly across legislative districts. By varying the way in which those profits enter the welfare function of the median legislator (before lobbying takes place), I can provide results about how district composition affects trade policy.

@@

I also plan to use the richness of this model to parse out how much of the changes in tariffs comes from the fact that the executives place lower weight on producer profits and how much comes from the fact that they make an agreement---that is, compare the effects of structure versus preferences.
\end{comment}

@@

Without making stronger assumptions, it is not possible to say how much the effort of each lobby will be reduced therefore whether total lobbying effort rises or falls. What is clear is that the optimization of the executives becomes significantly more complex as they are now faced with the problem of how to best exploit the free-riding dynamics that occur between the lobbies.

>>>>

I will begin by describing the model in detail. Section~\ref{sec:main} then presents the main results. In Section~ \ref{sec:example}, an extended example demonstrates these results as well as the impact of changing the institutional environment. Section~\ref{sec:dis} explores the connection between the model under consideration here and the `Protection for Sale' theoretical framework and includes in-depth discussions of the roles of uncertainty and the separation-of-powers structure. Section~\ref{sec:concl} concludes.

@@

I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small. This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}

@@

\subsection{The Trade Agreement}
\label{sec:ta}
The executives choose trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ to maximize Equation \ref{eq:jv}. I call the maximized joint value $MV(\bta)$ and write the division of surplus according to the Nash bargaining solution, which, due to symmetry, can be simplified from its general multiplicative form as
\[
  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
\]
\[
  V_E^*(\bta) - t = W_E^*(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
\]
where expectation operators have been dropped for convenience and $t$ is an intergovernmental transfer.\footnote{While direct monetary transfers have to date rarely been used in practice, it seems appropriate to interpret linked concessions on non-trade issues as indirect transfers (Klimenko, Ramey and Watson (2008); Maggi and Staiger (2011)). Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in asymmetric environments.} The solution to this system of equations determines the tariffs $\bta$ that will be set in the trade agreement.

@@

The problem then becomes to maximize the following modified version of Equation \ref{eq:jv}:
\begin{equation}
    \left\{ o \cdot B(\bta) + o^* \cdot B^*(\bta) \right\} \bm{W_E}(\btn) + \left\{ 1- o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \bm{W_E}(\bta)  
  \label{eq:jv2}
\end{equation}
where I have written the sum of the home and foreign executives' utilities as $\bm{W_E}(\cdot)$. Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement. One can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.

@@

\begin{comment}
\section{Extensions}
\label{sec:ext}

This analysis has been undertaken with the assumption that the import-competing industry is distributed evenly across legislative districts. By varying the way in which those profits enter the welfare function of the median legislator (before lobbying takes place), I can provide results about how district composition affects trade policy.

@@

I also plan to use the richness of this model to parse out how much of the changes in tariffs comes from the fact that the executives place lower weight on producer profits and how much comes from the fact that they make an agreement---that is, compare the effects of structure versus preferences.

@@

Without making stronger assumptions, it is not possible to say how much the effort of each lobby will be reduced therefore whether total lobbying effort rises or falls. What is clear is that the optimization of the executives becomes significantly more complex as they are now faced with the problem of how to best exploit the free-riding dynamics that occur between the lobbies. 
\end{comment}
